---
title: Environmental Analysis - How do sites differ and does this determine rates
  of clonality?
author: "Amanda Stahlke"
date: "5/13/2020"
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(BiodiversityR)
library(vegan)
library(goeveg)
library(lme4)
library(MuMIn)
library(rpart)
library(party)
library(subselect)
library(MASS)

library(dplyr)
library(reshape2)
library(ggplot2)
```

# Import data and subset

```{r prepdata}
gen_env<-na.omit(read.csv("geno_enviro.csv"))
names(gen_env)
#gen_env$BL<-gen_env$bare+gen_env$litter #composite background variable
gen_env$Site<-as.factor(gen_env$Site)
head(gen_env)
```

```{r GNrank}
site_GN_rank <- gen_env %>%
  group_by(Site) %>%
  summarise(unique(GN_Rank))
site_GN_rank 

clone_rate <- gen_env %>%
  group_by(Site) %>%
  summarise(sum(Clone)) %>%
  mutate(clone_rate=1-`sum(Clone)`/15) %>%
  as.data.frame()
merge(clone_rate, site_GN_rank)

stemcts <- gen_env %>%
  # group_by(Site) %>%
  summarise(avg=mean(stem), sd=sd(stem)) %>%
  mutate(avg/(100))

arrange(gen_env, stem) %>%
  dplyr::select(site.plot.subplot, stem)

```

## Create vegetation matrix
Aggregate site variables for the average and standard deviation across a site.

```{r aggregate_and_GNrank}
#####Veg summary
veg_mat<-aggregate(list("grass"=gen_env$grass,"bare"=gen_env$bare,"draba"=gen_env$Ldraba.cvr,"litter"=gen_env$litter),
                   by=list("Site"=gen_env$Site),FUN=mean)

veg_mat.sd<-aggregate(list("grass"=gen_env$grass,"bare"=gen_env$bare,"draba"=gen_env$Ldraba.cvr,"litter"=gen_env$litter),
                   by=list("Site"=gen_env$Site),FUN=sd)

GN<-read.csv("G_N_bysite.csv")
veg_mat$R_GN<-GN[,"R_GN"]
veg_mat$GN<-GN$site.G.N
veg_mat
```

R_GN is the rank of G/N across sites.
From least to most clonal: Site 1, 6, 3, 5, 2, 4. 
For the following plots, these variables are plotted against the G/N rank. 
I'm not sure how I feel about this because GN isn't uniformly distributed. Sites 3 and 6 have almost the same GN, for example.


```{r vegplotsgrass}
plot(veg_mat$R_GN,veg_mat$grass)
plot(veg_mat$R_GN,veg_mat$bare)
plot(veg_mat$R_GN,veg_mat$litter)
plot(veg_mat$R_GN,veg_mat$draba,xaxt="n",ylim=c(0,60),ylab="% Draba",xlab="Prevalence of Clones")
plot(veg_mat$R_GN,veg_mat$grass,xaxt="n",ylim=c(0,60),ylab="% Grass", xlab="Prevalence of Clones")
plot(veg_mat$R_GN,veg_mat$bare,xaxt="n",ylim=c(0,60),ylab="% Bare",xlab="Prevalence of Clones")
```

Grass is highly variable across sites. Bare ground seems to best seperate the low and high clonal sites.The most clonal sites have the most bare ground (but it's still only about 10%).

```{r}
 ggplot(data = gen_env, aes(x=Site, y=litter)) +
  geom_boxplot() + 
  geom_point(alpha=0.3)
 ggplot(data = gen_env, aes(x=Site, y=grass)) +
  geom_boxplot() + 
  geom_point(alpha=0.3)
 ggsave("grass_site.jpg")
 ggplot(data = gen_env, aes(x=Site, y=bare)) +
  geom_boxplot() + 
  geom_point(alpha=0.3)
  ggsave("bare_site.jpg")
 ggplot(data = gen_env, aes(x=Site, y=stem)) +
  geom_boxplot() + 
  geom_point(alpha=0.3) 
   ggsave("stem_site.jpg")
ggplot(data = gen_env, aes(x=Site, y=Ldraba.cvr)) +
  geom_boxplot() + 
  geom_point(alpha=0.3) 
   ggsave("Ldraba.cvr_site.jpg")
   
```


# Correlation among vriables 

Next, look at soil nutrient variables. 
```{r correlation}
cormat <- round(cor(gen_env[,c(7:9,11:dim(gen_env)[2])],  method = "spearman"),2) # drop texture and site

# Get upper triangle of the correlation matrix
  get_upper_tri <- function(cormat){
    cormat[lower.tri(cormat)]<- NA
    return(cormat)
  }
  
upper_tri <- get_upper_tri(cormat)
melted_cormat <- melt(upper_tri, na.rm = TRUE)

ggplot(data = melted_cormat, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
   name="Spearman\nCorrelation") +
  theme_minimal()+ 
 theme(axis.text.x = element_text(angle = 45, vjust = 1, 
    hjust = 1))+
 coord_fixed()

melted_cormat %>%
  filter(abs(value) >.5 & value <1) %>%
  arrange(desc(value))

melted_cormat %>%
  filter(Var1 == "Clone" |  Var2 == "Clone") %>%
  arrange(desc(value))

melted_cormat %>%
  filter(Var1 == "stem" |  Var2 == "stem") %>%
  arrange(desc(abs(value)))

melted_cormat %>%
  filter(Var1 == "bare" |  Var2 == "bare") %>%
  arrange(desc(abs(value)))

```

## Drop highly correlated variables for envfact

```{r drop_correlation_envfact}
names(gen_env)
env_fact<-gen_env[,11:dim(gen_env)[2]]
names(env_fact)

fact_drops<-c("Ca","Clay","Cu","Salts")
env_fact<-data.frame(env_fact[,-which(names(env_fact)%in%fact_drops)])
head(env_fact)

cormat <- round(cor(env_fact),2)
reshape2::melt(cormat) %>%
  filter(abs(value) >.5 & value <1)
```

Drop Ca, Clay, Cu, and Salts

#NMDS of subplots
```{r NMDS}
env.mds<-metaMDS(env_fact, distance = "gower", k=4)
stressplot(env.mds)
```

```{r NMDS2}
env.mds2<-metaMDS(env_fact, distance = "gower", k=4, previous.best = env.mds)
stressplot(env.mds2)
```

*** Solution reached
Non-metric fit is better than linear 

```{r stress_dims}
dim.result <- dimcheckMDS(env_fact,distance="gower",k=8,trymax = 20)
data.frame(dimension = seq_along(dim.result), stress=dim.result)
```
Stress is still barely above .2 at 2 dimensions, might as well show 3 dims. 

```{r NMDS_plot_site, warning=F, message=F}
env.plot<-ordiplot(env.mds,display = "sites",type="n", choices = c(1,2))

ordisymbol(env.plot, gen_env, "Site", legend=TRUE,cex=2,lwd=2,choices = c(1,2))
env.factors<-envfit(env.mds,env_fact, permutations = 999)
plot(env.factors, p.max=0.01)
```

Here, symbols represent the G/N ranking, where 1 is fewest clones and 6 is most. 
```{r NMDS_plot_GNrank, warning=F, message=F}
env.plot<-ordiplot(env.mds,display = "sites",type="n", choices = c(1,2))

ordisymbol(env.plot, gen_env, "GN_Rank", legend=TRUE,cex=2,lwd=2,choices = c(1,2))
env.factors<-envfit(env.mds,env_fact)
plot(env.factors)
```

# Examine multivariate dispersion among sites
Betadispersion - sites. Create a dissimilarity matrix for betdispersion analysis; Looks at multivariate dispersion (variance) among SITES

```{r betadisp}
commax<-vegdist(env_fact, method = "gower")
fum<-with(gen_env,betadisper(commax,Site))
plot(fum)
boxplot(fum) #variation among sampled points
permutest(fum) #-- Difference in multivariate dispersion between sites
TukeyHSD(fum)
```

# Make composite soil variables 
## from soils NMDS

```{r make_soilz_MDS}
soilz<-gen_env[,c("pH","OM","Nitrate","K","S",
                  "Zn","Fe","Mn","Mg","Na","CEC","P")] 

soilz.mds<-metaMDS(soilz,distance = "gower", k=4)
soilz.mds<-metaMDS(soilz,distance = "gower", k=4,
                   previous.best = soilz.mds)

dim.soil.result <- dimcheckMDS(soilz,distance="gower",k=4,trymax = 20)
data.frame(dimension = seq_along(dim.soil.result), stress=dim.soil.result)

## soil variables for multimodel comparison
# soil.nut1<-soilz.mds$points[,"MDS1"]
# soil.nut2<-soilz.mds$points[,"MDS2"]
```


```{r plot_soilz_NMDS}
nut.plot<-ordiplot(soilz.mds,display = "sites",type="n")
ordisymbol(nut.plot, gen_env, "Site", legend=TRUE,cex=2,lwd=2)
nut.factors<-envfit(soilz.mds,soilz)
plot(nut.factors)

```

## Make soil variables from PCoA

```{r soilz_pcoa}
soilz.D <- vegdist(soilz, "gower")
res <- ape::pcoa(soilz.D)
nut.factors<-envfit(soilz.D, soilz, perm=999)

biplot(res, rn=gen_env$Site)
plot(nut.factors, p.max=.05)

#soil variables for multimodel comparison
soil.nut.pcoa1<-res$vectors[,"Axis.1"]
soil.nut.pcoa2<-res$vectors[,"Axis.2"]

barplot(res$values$Relative_eig)
cumsum(res$values$Relative_eig)
cumsum(res$eig)/sum(res$eig)

```
# Do we need to account for site? 
```{r anosim_adonis, cache=TRUE}
env2 <- gen_env[,c(11:ncol(gen_env))]
env2 <- as.matrix(env2)
colnames(env2)

# vegdist(env2, distance = "gower")
anosim <- anosim(env2, grouping = gen_env$Site, distance = "gower", permutations = 9999) 
anosim
plot(anosim)

adonis <- adonis2(env2 ~ gen_env$Site)
adonis
```


# Global model for multi-model comparison; Clone 0,1

glm() fits a generalized linear model and glmer() fits linear mixed effects models
Global model includes Sand, grass, litter, soil, poca nutrients 1, pcoa nutrients 2. glmer allows inclusions of site as a random factor. 

```{r}
cl_global_bare_glm <- glm(Clone ~ Sand+grass+bare+
                       soil.nut.pcoa1+soil.nut.pcoa2,
                    data=gen_env,family="binomial",
                    na.action = "na.fail",weights = rep(15,dim(gen_env)[1]))
summary(cl_global_bare_glm)

cl_global_litter_glm <- glm(Clone ~ Sand+grass+litter+
                       soil.nut.pcoa1+soil.nut.pcoa2,
                    data=gen_env,family="binomial",
                    na.action = "na.fail",weights = rep(15,dim(gen_env)[1]))
summary(cl_global_litter_glm)

cl_global_bare_litter_site <- glmer(Clone ~ Sand+grass+bare+litter+
                        soil.nut.pcoa1 + soil.nut.pcoa2
                     +(1|gen_env$Site),
                     data=gen_env, family="binomial", 
                     na.action = "na.fail",
                     weights = rep(15,dim(gen_env)[1]))
summary(cl_global_bare_litter_site)

cl_global_bare_site <- glmer(Clone ~ Sand+grass+bare+
                        soil.nut.pcoa1 + soil.nut.pcoa2
                     +(1|gen_env$Site),
                     data=gen_env, family="binomial", 
                     na.action = "na.fail",
                     weights = rep(15,dim(gen_env)[1]))
summary(cl_global_bare_site)

cl_global_litter_site <- glmer(Clone ~ Sand+grass+litter+
                        soil.nut.pcoa1 + soil.nut.pcoa2
                     +(1|Site),
                     data=gen_env, family="binomial", 
                     na.action = "na.fail",
                     weights = rep(15,dim(gen_env)[1]))
summary(cl_global_litter_site)

anova( cl_global_litter_site, cl_global_litter_glm)
anova(cl_global_bare_site, cl_global_litter_site)

cl_redux1_litter_site <- glmer(Clone ~ Sand+grass+litter+
                        soil.nut.pcoa1 + 
                     +(1|Site),
                     data=gen_env, family="binomial", 
                     na.action = "na.fail",
                     weights = rep(15,dim(gen_env)[1]))
summary(cl_redux1_litter_site)

anova( cl_global_litter_site, cl_redux1_litter_site)


cl_redux2_litter_site <- glmer(Clone ~ Sand+litter+
                        soil.nut.pcoa1 + 
                     (1|Site),
                     data=gen_env, family="binomial", 
                     na.action = "na.fail",
                     weights = rep(15,dim(gen_env)[1]))
summary(cl_redux2_litter_site)

anova( cl_redux1_litter_site, cl_redux2_litter_site)


```

Inclusion of site as random effect impoves AIC. The original model included bare greound. When site was included as a random effect, bare ground lots significance. Not suprising since most subplots did not have any bareground except for at Site 4 which had the most clones. 
When site is included as a random effect, sand gains significance. 

Any points with extra high influence of the glm? Can't run Cook's distance with the mixed effects model unfortunately. 


```{r}
cooksd<-cooks.distance(cl_global_litter_glm)
plot(cooksd)
abline(h = 4*mean(cooksd, na.rm=T), col="red")
text(x=1:length(cooksd)+1, y=cooksd, labels=ifelse(cooksd>4*mean(cooksd, na.rm=T),names(cooksd),""), col="red")  # add labels
influential <- as.numeric(names(cooksd)[(cooksd > 4*mean(cooksd, na.rm=T))])

cbind(gen_env[influential, ],soil.nut.pcoa1[influential],soil.nut.pcoa2[influential])

```
Remove these subplots and run again.

Tried testing for highly-influential points in the mixed effects model with influence.ME package -
https://journal.r-project.org/archive/2012-2/RJournal_2012-2_Nieuwenhuis~et~al.pdf
but could not resolve annoying lengths differ error. 
"Error in model.frame.default(data = data.update, weights = rep(15, dim(gen_env)[1]), : variable lengths differ (found for 'soil.nut.pcoa1')"

```{r influence.ME}
#install.packages("influence.ME")
# library(influence.ME)
# estex <- influence.ME::influence(cl_global_litter_site, "Site")
```

```{r}
gen_env_pruned <- gen_env[-influential,]
soil.nut.pcoa1_pruned<-res$vectors[-influential,"Axis.1"]
soil.nut.pcoa2_pruned<-res$vectors[-influential,"Axis.2"]

cl_global_glm_pruned<-glm(Clone ~ 
           Sand+grass+litter+
             soil.nut.pcoa1_pruned+soil.nut.pcoa2_pruned,
           data=gen_env_pruned, family="binomial",
           na.action="na.fail",
           weights = rep(13,dim(gen_env_pruned)[1]))

cl_global_glmer_pruned<-glmer(Clone ~ 
           Sand+grass+litter+
             soil.nut.pcoa1_pruned + soil.nut.pcoa2_pruned 
           +(1|gen_env_pruned$Site),
           data=gen_env_pruned, 
           family="binomial", 
           na.action = "na.fail",
           weights = rep(13,dim(gen_env_pruned)[1]))

summary(cl_global_glm_pruned)
summary(cl_global_glmer_pruned)
```



```{r}
cooksd_pruned<-cooks.distance(cl_global_glm_pruned)
plot(cooksd_pruned)
abline(h = 4*mean(cooksd_pruned, na.rm=T), col="red")
text(x=1:length(cooksd_pruned)+1, y=cooksd_pruned,
     labels=ifelse(cooksd_pruned > 4*mean(cooksd_pruned,na.rm=T),
                   names(cooksd_pruned),""), col="red")  # add labels
influential <- as.numeric(names(cooksd_pruned)
                          [(cooksd_pruned > 4*mean(cooksd_pruned,
                                                   na.rm=T))])
cbind(gen_env_pruned[influential,],
      soil.nut.pcoa1_pruned[influential],
      soil.nut.pcoa2_pruned[influential])
```

## Does site 4 have an outsized effect? 

```{r drop_4}
drop4 <- which(gen_env$Site == 4)
gen_env_pruned4 <- gen_env[-drop4,]
soil.nut.pcoa1_pruned4 <-res$vectors[-drop4,"Axis.1"]
soil.nut.pcoa2_pruned4 <-res$vectors[-drop4,"Axis.2"]


cl_global_glm_pruned4<-glm(Clone ~
                            Sand+grass+litter+
                            soil.nut.pcoa1_pruned4 + soil.nut.pcoa2_pruned4,
                          data=gen_env_pruned4,
                          family="binomial",
                          na.action="na.fail",
                          weights = rep(15,dim(gen_env_pruned4)[1]))


cl_global_glmer_pruned4<-glmer(Clone ~
                                Sand+grass+litter+
                                soil.nut.pcoa1_pruned4 + soil.nut.pcoa2_pruned4
                              +(1|gen_env_pruned4$Site),
                              data=gen_env_pruned4,
                              family="binomial",
                              na.action = "na.fail",
                              weights = rep(15,dim(gen_env_pruned4)[1]))

summary(cl_global_glm_pruned4)
summary(cl_global_glmer_pruned4)

anova(cl_global_glmer_pruned4, cl_global_glm_pruned4)
```
After dropping site 4, the glmer incorporating site was still a significantly better model.The glm identified the intercept, Sand, bare, and soil.nut1 as highly significant.  The glmer identfied Sand, grass, bare*, and soil.nut1. 

Outliers? 

```{r}
cooksd_pruned <- cooks.distance(cl_global_glm_pruned4)
plot(cooksd_pruned)
abline(h = 4*mean(cooksd_pruned, na.rm=T), col="red")
text(x=1:length(cooksd_pruned)+1, y=cooksd_pruned,
     labels=ifelse(cooksd_pruned > 4*mean(cooksd_pruned,na.rm=T),
                   names(cooksd_pruned),""), col="red")  # add labels
influential <- as.numeric(names(cooksd_pruned)
                          [(cooksd_pruned > 4*mean(cooksd_pruned, na.rm=T) )] )

cbind(gen_env_pruned4[influential,],
      soil.nut.pcoa1_pruned4[influential],
      soil.nut.pcoa1_pruned4[influential])

```
As before, 2.3.11 and 6.1.5 are outliers. Leave these in and more on. Continue to parse down the glmer without site 4. 

redux 1 - drop soil.nut2; not significant
redux 2 - drop bare; less significant
Significantly improved AIC.
```{r}
cl_redux1_glmer_pruned4<-glmer(Clone ~
                                Sand+grass+bare+
                                soil.nut.pcoa1_pruned4
                              +(1|gen_env_pruned4$Site),
                              data=gen_env_pruned4,
                              family="binomial",
                              na.action = "na.fail",
                              weights = rep(15,dim(gen_env_pruned4)[1]))
summary(cl_redux1_glmer_pruned4)

cl_redux1_glmer_pruned4<-glmer(Clone ~
                                Sand+litter+grass+
                                soil.nut.pcoa1_pruned4
                              +(1|gen_env_pruned4$Site),
                              data=gen_env_pruned4,
                              family="binomial",
                              na.action = "na.fail",
                              weights = rep(15,dim(gen_env_pruned4)[1]))
summary(cl_redux1_glmer_pruned4)

cl_redux2_glmer_pruned4<-glmer(Clone ~
                                Sand+litter+
                                soil.nut.pcoa1_pruned4
                              +(1|gen_env_pruned4$Site),
                              data=gen_env_pruned4,
                              family="binomial",
                              na.action = "na.fail",
                              weights = rep(15,dim(gen_env_pruned4)[1]))
summary(cl_redux2_glmer_pruned4)

anova(cl_global_glmer_pruned4, cl_redux2_glmer_pruned4)

cl_redux2_glmer_pruned4<-glmer(Clone ~
                                Sand+grass+
                                soil.nut.pcoa1_pruned4
                              +(1|gen_env_pruned4$Site),
                              data=gen_env_pruned4,
                              family="binomial",
                              na.action = "na.fail",
                              weights = rep(15,dim(gen_env_pruned4)[1]))
summary(cl_redux2_glmer_pruned4)


```

With site 4, reduce model. 

```{r}
cl_redux1_glmer_pruned<- glmer(Clone ~ 
           Sand+ grass + bare +
             soil.nut.pcoa1_pruned +
           +(1|gen_env_pruned$Site),
           data=gen_env_pruned, 
           family="binomial", 
           na.action = "na.fail",
           weights = rep(13,dim(gen_env_pruned)[1]))
summary(cl_redux1_glmer_pruned)

anova(cl_global_glmer_pruned, cl_redux1_glmer_pruned)

cl_redux2_glmer_pruned <- glmer(Clone ~ 
            Sand+ grass
              + soil.nut.pcoa1_pruned +
              (1|gen_env_pruned$Site),
             data=gen_env_pruned,family="binomial",
            na.action = "na.fail",
             weights = rep(13,dim(gen_env_pruned)[1]))
summary(cl_redux2_glmer_pruned)

anova(cl_global_glmer_pruned, cl_redux2_glmer_pruned)

cl_redux3_glmer_pruned <- glmer(Clone ~ 
            Sand + soil.nut.pcoa1_pruned +
              (1|gen_env_pruned$Site),
             data=gen_env_pruned,family="binomial",
            na.action = "na.fail",
             weights = rep(13,dim(gen_env_pruned)[1]))
summary(cl_redux3_glmer_pruned)

cl_redux4_glmer_pruned <- glmer(Clone ~ 
            grass + soil.nut.pcoa1_pruned +
              (1|gen_env_pruned$Site),
             data=gen_env_pruned,family="binomial",
            na.action = "na.fail",
             weights = rep(13,dim(gen_env_pruned)[1]))

anova(cl_redux2_glmer_pruned, cl_redux3_glmer_pruned)
```

## Can we identify individual soil variables that predict subplot clonality? 

```{r}
cl_redux_nutrs_glmer_pruned <- glmer(Clone ~
              Sand + litter + grass +                       
              I(K/100)+ OM + Fe + Zn + S + Mn
              +  (1|gen_env_pruned$Site),
             data=gen_env_pruned,family="binomial",
            na.action = "na.fail",
             weights = rep(13,dim(gen_env_pruned)[1]))
summary(cl_redux_nutrs_glmer_pruned)

dd<-dredge(cl_redux_nutrs_glmer_pruned)
subset(dd,delta<4)
importance(dd)#)
importance(subset(dd, delta<=2))


## without site 4
cl_redux_nutrs_glmer_pruned4 <- glmer(Clone ~
                              grass + Sand +                        
                              I(K/100) + OM + Fe + Zn + S + Mn
                              + (1|gen_env_pruned4$Site),
                              data=gen_env_pruned4,
                              family="binomial",
                              na.action = "na.fail",
                              weights = rep(15,dim(gen_env_pruned4)[1]))
summary(cl_redux_nutrs_glmer_pruned4)

dd4<-dredge(cl_redux_nutrs_glmer_pruned4)
subset(dd4,delta<4)
importance(dd4)#)
importance(subset(dd4, delta<=2))

?dredge

```


## Try re-scaling soil.nutrients or grass/sand/bare?

```{r rescale_vars}
max(soil.nut.pcoa1_pruned4) - min(soil.nut.pcoa1_pruned4)
max(soil.nut.pcoa2_pruned4) - min(soil.nut.pcoa2_pruned4)

max(soil.nut.pcoa1_pruned) - min(soil.nut.pcoa1_pruned)
max(gen_env_pruned$Sand) - min(gen_env_pruned$Sand)
max(gen_env_pruned$grass) - min(gen_env_pruned$grass)

max(gen_env_pruned$K/100) - min(gen_env_pruned$K/100)
max(gen_env_pruned$OM) - min(gen_env_pruned$OM)

cl_global_glmer_pruned_rescaled <-glmer(Clone ~ 
           I(Sand/100) + I(grass/100) +I(litter/100)+
             soil.nut.pcoa1_pruned +
             soil.nut.pcoa2_pruned                           
           +(1|gen_env_pruned$Site),
           data=gen_env_pruned, 
           family="binomial", 
           na.action = "na.fail",
           weights = rep(13,dim(gen_env_pruned)[1]))
summary(cl_global_glmer_pruned_rescaled)


cl_global_glmer_pruned_rescaled2 <-glmer(Clone ~ 
           I(Sand/100) + I(grass/100) +I(litter/100)+
             soil.nut.pcoa1_pruned +
           +(1|gen_env_pruned$Site),
           data=gen_env_pruned, 
           family="binomial", 
           na.action = "na.fail",
           weights = rep(13,dim(gen_env_pruned)[1]))
summary(cl_global_glmer_pruned_rescaled2)


cl_redux_nutrs_glmer_pruned4_rescale<-glmer(Clone ~
                                      I(Sand/100) + 
                                        I(grass/100) + 
                                        I(litter/100) +
                              I(K/100)+ OM + Fe + Zn + S + Mn
                              +(1|gen_env_pruned$Site),
                              data=gen_env_pruned,
                              family="binomial",
                              na.action = "na.fail",
                              weights = rep(15,dim(gen_env_pruned)[1]))
summary(cl_redux_nutrs_glmer_pruned4_rescale)

dd2_4<-dredge(cl_redux_nutrs_glmer_pruned4_rescale)
subset(dd2_4,delta<=2)
importance(dd2_4)#)
importance(subset(dd2_4, delta<4))


cl_redux_nutrs_glmer_pruned_rescale<-glmer(Clone ~
                                      I(Sand/100) + 
                                        I(grass/100) +I(litter/100)+
                              I(K/100)+ OM + Fe + Zn + S + Mn
                              +(1|gen_env_pruned$Site),
                              data=gen_env_pruned,
                              family="binomial",
                              na.action = "na.fail",
                              weights = rep(13,dim(gen_env_pruned)[1]))
summary(cl_redux_nutrs_glmer_pruned_rescale)


dd2<-dredge(cl_redux_nutrs_glmer_pruned_rescale)
subset(dd2,delta<4)
importance(dd2)#)
importance(subset(dd2, delta<4))


cl_redux2_nutrs_glmer_pruned_rescale<-glmer(Clone ~
                                      I(Sand/100) + 
                                        I(grass/100) +
                                        I(litter/100)+
                               OM + S 
                              +(1|gen_env_pruned$Site),
                              data=gen_env_pruned,
                              family="binomial",
                              na.action = "na.fail",
                              weights = 
                                rep(13, dim(gen_env_pruned)[1]))
summary(cl_redux2_nutrs_glmer_pruned_rescale)

## Taking out site 4 -> lose significance of grass
cl_redux2_nutrs_glmer_pruned4_rescale<-glmer(Clone ~
                                      I(Sand/100) + 
                                        I(grass/100) +
                                        I(litter/100)+
                               OM + S 
                              +(1|gen_env_pruned4$Site),
                              data=gen_env_pruned4,
                              family="binomial",
                              na.action = "na.fail",
                              weights = 
                                rep(13, dim(gen_env_pruned4)[1]))
summary(cl_redux2_nutrs_glmer_pruned4_rescale)

```

OM is highlighy negatively correlated with bare ground. (Check OM variation too.)

## Plot model results
```{r ggcoefstats}
library(ggstatsplot)

# for reproducibility
set.seed(123)


# plot
ggstatsplot::ggcoefstats(
  cl_redux2_nutrs_glmer_pruned_rescale,
  point.args = list(color = "red", shape = 15),
  vline.args = list(color = "#CC79A7", linetype = "dotdash"),
  # stats.label.color = c("#0072B2", "#D55E00", "darkgreen"),
  title = "Subplot clonality predicted by environmental conditions",
  subtitle = "Final GLMER with site as random effect",
  # ggtheme = hrbrthemes::theme_ipsum_ps(),
  ggstatsplot.layer = FALSE
) + # note the order in which the labels are entered
  # ggplot2::scale_y_discrete(labels = c("transmission", "cylinders", "interaction")) +
  ggplot2::labs(x = "regression coefficient", y = NULL)
  ggsave("subplotclonality_glmer_final.jpg")
dev.off()
```

# Do the same predictors of clonality predict Stem density (1-inf)?
lmer for continuous cariable: count of stems with random effect of site. 

```{r}
stem_global_lmer_pruned_rescaled <-lmer(stem ~ 
           I(Sand/100) + I(grass/100) +I(litter/100)+
             soil.nut.pcoa1_pruned +
             soil.nut.pcoa2_pruned                           
           +(1|gen_env_pruned$Site),
           data=gen_env_pruned, 
           na.action = "na.fail")
summary(stem_global_lmer_pruned_rescaled)

stem_global_lm_pruned_rescaled <- lm(stem~
                     I(Sand/100) + I(grass/100) +I(litter/100)+
                      soil.nut.pcoa1_pruned +
                    soil.nut.pcoa2_pruned,
                      data=gen_env_pruned)
summary(stem_global_lm_pruned_rescaled)

anova(stem_global_lmer_pruned_rescaled, stem_global_lm_pruned_rescaled)

## Cook's distance?

stem_redux1_lm_pruned_rescaled <- lm(stem~
                     I(Sand/100) + I(grass/100) +I(litter/100)+
                    soil.nut.pcoa2_pruned,
                      data=gen_env_pruned)
summary(stem_redux1_lm_pruned_rescaled)

anova(stem_global_lm_pruned_rescaled, stem_redux1_lm_pruned_rescaled)

# Dropping site 4 does not make a difference.
# stem_global_lm_pruned_rescaled4 <- lm(stem~
#                      I(Sand/100) + I(grass/100) +I(litter/100)+
#                       soil.nut.pcoa1_pruned4 +
#                     soil.nut.pcoa2_pruned4,
#                       data=gen_env_pruned4)
# summary(stem_global_lm_pruned_rescaled4)

stem_nutrs_lm <- lm(stem ~
              I(Sand/100) + I(grass/100) +I(litter/100)+   
              I(K/100)+ OM + Fe + Zn + S + Mn,
              na.action = "na.fail",
             data=gen_env)
summary(stem_nutrs_lm)

dd<-dredge(stem_nutrs_lm)
subset(dd,delta<=2)
importance(dd)#)
importance(subset(dd, delta<=2))

stem_redux_nutrs_lm <- lm(stem ~
              I(Sand/100) + I(grass/100) +I(litter/100)+   
              OM,
              na.action = "na.fail",
             data=gen_env)
summary(stem_redux_nutrs_lm)

ggstatsplot::ggcoefstats(
  stem_redux_nutrs_lm,
  point.args = list(color = "red", shape = 15),
  vline.args = list(color = "#CC79A7", linetype = "dotdash"),
  # stats.label.color = c("#0072B2", "#D55E00", "darkgreen"),
  title = "Subplot whitetop stem density predicted by environmental conditions",
  subtitle = "Final LM ",
  # ggtheme = hrbrthemes::theme_ipsum_ps(),
  ggstatsplot.layer = FALSE
) + # note the order in which the labels are entered
  # ggplot2::scale_y_discrete(labels = c("transmission", "cylinders", "interaction")) +
  ggplot2::labs(x = "regression coefficient", y = NULL)
  ggsave("subplot_stemdensity_lm_final.jpg")
dev.off()

```


paring down the soil data
have to NAomit as missing NH4_N values for rows 1,18,and 32

```{r}
soil<-na.omit(gen_env[,c("Site","pH","OM","Nitrate","K","S","Zn","Fe","Mn","Mg","Na","CEC","P")])
irisHmat <- ldaHmat(soil[,2:dim(soil)[2]],soil$Site)
eleaps(irisHmat$mat,kmin=2)#,H=irisHmat$H,r=irisHmat$r,crit="ccr12")
```

```{r}
fit <- lda(Site ~ K + Mg + Na, data=soil, na.action=na.exclude,
           CV=TRUE)
fit # show results
```


# Canonical Analysis of Principal Coordinates 
As in Andreson & Willis 2003, Biodiversity R package
In constrained ordination we do not want to display all or even most ofthe compositional variation, but only the variation that can be explained by the used environmental variables, or constraints.

```{r}
gen_env$factClone<-as.factor(gen_env$Clone)

Ordination.model1 <- CAPdiscrim(soilz~factClone,
                                data=gen_env,
                                dist="gower", 
                                axes=3, m=0, add=FALSE)
Ordination.model1

# data(env_fact)
# Ordination.model1 <- CAPdiscrim(env_fact~Site,
#                                 data=gen_env,
#                                 dist="gower",
#                                 axes=2, m=0,
#                                 add=FALSE,
#                                 permutations = 100)
# Ordination.model1

plot1 <- ordiplot(Ordination.model1)
ordisymbol(plot1, gen_env, "Clone", legend=TRUE)

sitz<-envfit(Ordination.model1, 
             env_fact[,c("grass","OM","S","Sand","Fe","Zn")], 
             choices = c(1,2))
sitz.text<-envfit(Ordination.model1,gen_env[,c("bare","grass","Mg","Na","K")])
ordisymbol(plot1, gen_env, "Site", legend=TRUE)
plot(sitz)
plot(sitz.text)
plot2 <- ordiplot(Ordination.model1, type="none",choices = c(1,2))
ordisymbol(plot2, gen_env,"Site", legend = T)
plot(sitz)

```



```{r}
# plot change in classification success against m (#classification dimensions for LDA)
plot(seq(1:14), rep(-1000, 14), xlim=c(1, 14), ylim=c(0, 100), xlab="m",
     ylab="classification success (percent)", type="n")

for (mseq in 1:14) {
  CAPdiscrim.result <- CAPdiscrim(soilz~Site, data=gen_env, 
                                  dist="gower", axes=2, m=mseq)
  points(mseq, CAPdiscrim.result$percent)
}

```

```{r}
stMod.1<-glm(Clone~grass+bare+Zn+Fe+OM+Mn,data=gen_env[gen_env$Site==1,],family="binomial",na.action = "na.fail")
summary(stMod.1)
stepAIC(stMod.1,direction = "both")
```


```{r}
gen_env$factClone<-as.factor(gen_env$Clone)

env.mds2

#### use vegetation matrix instead
Ordination.model1 <- CAPdiscrim(soilz~factClone,
                                data=gen_env,
                                dist="gower", 
                                axes=3, m=0, add=FALSE)
Ordination.model1

```

```{r}
plot1 <- ordiplot(Ordination.model1, type="none")
# ordisymbol(plot1, gen_env, "Clone", legend=TRUE)
ordisymbol(plot1, gen_env, "Site", legend=TRUE)
clonz<-envfit(Ordination.model1,soilz)
plot(clonz)
```

Plot change in classification success against m (#classification dimensions for LDA)
```{r}
plot(seq(1:14), rep(-1000, 14), xlim=c(1, 14), ylim=c(0, 100), xlab="m", 
     ylab="classification success (percent)", type="n")
for (mseq in 1:14) {
  CAPdiscrim.result <- CAPdiscrim(soilz~Site, data=gen_env, 
                                  dist="gower", axes=2, m=mseq)
  points(mseq, CAPdiscrim.result$percent)
}


summary(CAPdiscrim(soilz~Site, data=gen_env, 
                                  dist="gower", axes=2))
```


```{r}
# #####Diversity and Evenness (as calculated in Gaskin 2006)
# bySite<-read.csv("G_N_bysite.csv")
# byPlot<-read.csv("G_N_bysiteplot.csv")
# GID_S<-read.csv("G_ID_Site.csv")
# GID_P<-read.csv("G_ID_Plot.csv")
# 
# GID_P$Site.Plot<-paste(GID_P$Site,GID_P$Plot,sep = ".")
# byPlot$Div<-0
# byPlot$Evn<-0
# 
# Di<-function(ram, N) { sum(ram*(ram-1)) / (N*(N-1))  }
# 
# 
# for(i in seq_along(byPlot$site.plot)) byPlot[i,"Div"]<- 1 - Di(GID_P[GID_P$Site.Plot==byPlot$site.plot[i],"N_G_ID"],10)
# 
# Dmin<-function(G,N) { ((G-1)*(2*N-G))/(N*(N-1)) }
# Dmax<-function(G,N) { (N*(G-1))/(G*(N-1)) }
# Ev<-function(D,Dmi,Dma) { (D - Dmi)/(Dma - Dmi)}
# 
# byPlot$Evn<-round(Ev(byPlot$Div,Dmi = Dmin(G=byPlot$plot.G,10),Dma = Dmax(byPlot$plot.G,10)),2)
#               
# plot(byPlot$Div,byPlot$Evn)  
# plot(byPlot$plot.freq,byPlot$Evn)
# plot(byPlot$av.stm.dens,byPlot$Evn)

```


# CART
Do the same variables that divide sites also classify clonality?
```{r}
fitSite <-rpart(Site ~ pH+OM+Nitrate+K+S+Zn+Fe+Mn+Mg+Na+CEC+P, method = "class", data = gen_env)
fitClone <-rpart(Clone ~ pH+OM+Nitrate+K+S+Zn+Fe+Mn+Mg+Na+CEC+P, method = "class", data = gen_env)
```

```{r}
printcp(fitSite)
plot(fitSite)
text(fitSite)
# plotcp(fitSite)
summary(fitSite)
```


```{r}
printcp(fitClone)
plotcp(fitClone)
summary(fitClone)
plot(fitClone)
text(fitClone)
```


```{r}
pfit<- prune(fitSite, cp=fitSite$cptable[which.min(fitSite$cptable[,"xerror"]),"CP"])
printcp(pfit)

plot(pfit, uniform=TRUE,
     main="Pruned Classification Tree for Site")
text(pfit, use.n=TRUE, all=TRUE, cex=.8)
```


```{r}
pfitClone<- prune(fitClone, cp=fitClone$cptable[which.min(fitClone$cptable[,"xerror"]),"CP"])
printcp(pfitClone)

plot(pfitClone, uniform=TRUE,
     main="Pruned Classification Tree for Clones")
text(pfitClone, use.n=TRUE, all=TRUE, cex=.8)

```

### CI tree
```{r}
fit <- ctree(Site ~ pH+OM+Nitrate+K+S+Zn+Fe+Mn+Mg+Na+CEC+P,
             data=gen_env)
plot(fit, main="Conditional Inference Tree for Site")
```


```{r}
######NMDS extras
# nut.labs<-c("K","Clay","Na","S","Mg","BareG","P")
# 
# 
# md.soilz<-envfit(soilz.mds,gen_env[,c("K","S","Mg","Na","P","Sand","bare")])
# # md.soilz<-envfit(soilz.mds,gen_env[,names(soilz)])
# plot(md.soilz,lwd=2,col="black")#,labels =" ")
# # 
# # nut.vars<-locator(7)
# # soilz.mds$points[,1:2]
# #nut.vars<-locator(7)
# nut.labs<-c("K","Clay","Na","S","Mg","BareG","P")
# text(nut.vars,nut.labs,cex=2)
# clonz<-read.csv("G_N_bysite.csv")
# clonz.loc<-locator(6)
# text(clonz.loc,1:6)

#library(stringi)
#byPlot$Site<-substr(byPlot$site.plot,1,1)
#byPlot$Plot<-substr(byPlot$site.plot,3,3)

#boxplot(byPlot$Evn~byPlot$Site)
#boxplot(byPlot$Div~byPlot$Site)

#bySite$Div<-0
#byPlot$Evn<-0

#for(i in seq_along(bySite$site)) bySite[i,"Div"]<- 1 - Di(GID_S[GID_S$Site==bySite$site[i],"N_G_ID"],30)
#bySite$Evn<-round(Ev(bySite$Div,Dmi = Dmin(G=bySite$site.G,30),Dma = Dmax(bySite$site.G,30)),2)
#plot(bySite$Div,bySite$Evn)


####Principal Components Analysis -- making a soil nutrient variable
#gen_env<-gen_env[-which(gen_env$site.plot.subplot=="4.2.6"),]

#soilvar<-princomp(~pH+Salts+OM+Nitrate+K+S+Zn+Fe+Mn+Cu+Mg+Na+CEC+P, 
#                  data=gen_env, cor=FALSE, scores=TRUE)

#soilvar<-princomp(~K+S+Mg+Na+P, 
#                  data=gen_env, cor=FALSE, scores=TRUE)

#summary(soilvar)
#screeplot(soilvar)

#loadings(soilvar)
#gen_env$soil.nut<-soilvar$scores[,1]
```